---
title: "BayesMultiLogit"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BayesMultiLogit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(BayesMultiLogit)
```

# BayesMultiLogit  

This is an R package developed by Kyle R. McEvoy and Jared Fisher to implement a number of methods for performing Bayesian multinomial logistic regression using data-augmentation methods. 

The code for all functions using the Holmes-Held methods were written by us following the pseudo-code template provided in the paper:
"Bayesian Auxiliary Variable Models for Binary and Multinomial Regression" by Chris C. Holmes and Leonhard Held, Bayesian Analysis (2006).

The code for the Polya-Gamma functions was copied from the public repository <https://github.com/jgscott/helloPG> by James G. Scott who was adapting code
from the package <https://github.com/jwindle/BayesLogit>. The BayesLogit repository includes code written by Jesse Windle, Nicholas Polson, and James G. Scott.

The multivariate normal density function was copied from work by Nino Hardt, Dicko Ahmadou, Benjamin Christoffersen on the Rcpp Gallery
located at <https://gallery.rcpp.org/articles/dmvnorm_arma/>.

## Using the Functions

The package contains five wrapper functions for performing Bayesian multinomial logistic regression. These wrappers validate dimensions, response values, iteration counts, flags, and prior parameters before calling the C++ samplers. `multilogit_AMH()` targets the multinomial likelihood directly; the remaining methods use data augmentation.

Each function expects matrix-like `Y` and `X` inputs. `Y` is an $N$ by $C$ matrix with one row per observation, and `X` is an $N$ by $P$ design matrix. The direct adaptive Metropolis-Hastings, augmented Metropolis-Hastings, augmented elliptical slice, and Polya-Gamma samplers accept non-negative category counts with a positive row total. The Holmes-Held sampler requires one-hot rows. The first column of `X` should be an intercept column of ones; centering and scaling the remaining columns is generally helpful.

When priors on the coefficients are specified the mean should be a vector with length equal to $P$ the number of columns of $X$ including the intercept column, and the prior variance should be a $P$ by $P$ square matrix. In general the functions do not allow for different priors across categories and they assume independence across categories as well.  

The C++ loops periodically check for user interrupts. As with any MCMC method, runtime depends on the data dimensions and requested number of iterations.

## Examples

```{r}

#Setting up data structures properly.

X <- scale(iris[, 1:4])

X <- cbind(1, X)

Y <- matrix(0, nrow = 150, ncol = 3)

Y <- sapply(c(1,2,3), function(x) Y[, x] <- as.numeric((as.numeric(iris$Species) == x) )) 

```

```{r}

#Running the wrapper functions:

out1 <- multilogit(Y, X, n_sample = 100, n_burn = 50, reference_cat = 1,
                   progress = FALSE)

out0 <- multilogit_AMH(Y, X, n_sample = 100, n_burn = 50,
                       reference_cat = 1, progress = FALSE)

out2 <- multilogit_ESS(Y, X, n_sample = 100, n_burn = 50,
                       probs = FALSE, progress = FALSE)

out3 <- multilogit_holmesheld(Y, X, v = diag(ncol(X)), n_sample = 100,
                              n_burn = 50, progress = FALSE)

out4 <- multilogit_PG(Y, X, n_sample = 100, n_burn = 50,
                      probs = TRUE, progress = FALSE)


```
## Outputs

Outputs are lists. `posterior_coef` is a $P$ by $C$ by `n_sample` array. When `probs = TRUE`, `posterior_prob` is an $N$ by $C$ by `n_sample` array.


```{r}
#Outputs

out1$posterior_coef[1, 1, 90:100]

out3$posterior_coef[3, 3, 90:100]
```

Time series plots are often useful for examining the chains. 

```{r}

plot.ts(out4$posterior_prob[113,3,])


```
